---
title: "Class 5"
output: html_notebook
---

##Chapter 5: Robust Functions

When we are working with functions we must be very clear when users make failures. We need to specify what is the error that he/she has commited displaying an informative message in the console.

```{r, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
# Define troublesome x and y
x <- c(NA, NA, NA)
y <- c( 1, NA, NA, NA)

both_na <- function(x, y) {
  # Replace condition with logical
  if (length(x)!=length(y)) {
    # Replace "Error" with better message
    stop("x and y must have the same length", call. = FALSE)
  }  
  
  sum(is.na(x) & is.na(y))
}

```

```{r, echo=TRUE, eval=FALSE, warning=FALSE, message=TRUE}
# Call both_na() 
both_na()
```

It shows a error message explaining that exists different input lenghts. For this reason we use `stop()` to be more precisely when we deal with failures. An important argument here is `call. = FALSE` that unlink the `Error` word of your error message. Otherwise, in `call.=TRUE`, the word `Error` goes along with your message.

### Unstable types



```{r}
df <- data.frame(
  a = 1L,
  b = 1.5,
  y = Sys.time(),
  z = ordered(1)
)


col_classes <- function(df) {
  class_list <- map(df, class)
  
  # Add a check that no element of class_list has length > 1
  #if (condition) {
  if (any(map_dbl(class_list, length)>1)) {
    stop("Some columns have more than one class", call. = FALSE)
  }
  
  # Use flatten_chr() to return a character vector
  flatten_chr(class_list)
}

# Check that our new function is type consistent
df %>% col_classes() %>% str()


# Check that our new function is type consistent
df %>% col_classes() %>% str()
df[3:4] %>% col_classes() %>% str()
df[1:2] %>% col_classes() %>% str()
```


