---
title: "Boost Statistical Computing through Functional Programming in R"
author: Omar Benites
output:
  html_document: 
    toc: yes
  html_notebook: default
---

```{r}
  knitr::clean_cache()

```

# Module 1: Introduction

Almost everything in R is based on functions or functional-approach programming. A function is a piece of code that we usually run in our analysis that are compresed in a expresed that can be re-used.

A function is compouned by:

- **Name of the function**: To put a function name you need the assignment operator  **<-**.  
- **The arguments** (formal argumentes) of the function. Could be whatever type of data structure.
- **The body**: for what the function is composed.
- **The environment**: it's the environment where the environment is defined. By the default is the global environment. This is imporntant because it is where the function looks for variables, this is called the *scope* of the function.


```{r, echo=TRUE, eval=TRUE}
add <- function(x,y = 1){
        x+y
}
##
formals(add)
body(add)
environment(add)
```

One important things that is the *return* command. If we use *return*, automatically the function answers with the result and then stop. For example


```{r, echo=TRUE, eval=TRUE}
minor <- function(x, y){
      r <- x-y #local variable r
      return(r)
      1+1
}

minor(x = 15, y =12)

```


Function are also objects. For example if we assign the function *mean* to new variable, it looks like:


```{r, echo=TRUE, eval=TRUE}
mean2 <- mean
v <- c(15, 12)
mean2(v)
```


# Environments 

## Scoping of R-functions

Scoping describes how R looks up values by name. For example

```{r, echo=TRUE, eval=TRUE}
f <- function() {
 x <- 1
 y <- 2
 c(x, y)
}
f()
```

In this case, the function *f* looks for the values which are inside the function. However exists another cases where the function looks for one level up. Let see:

```{r, echo=TRUE, eval=TRUE}
z <- 15
f <- function() {
 x <- 1
 y <- 2
 c(x, y, z)
}
f()
```
In this case, the function does not define **z** in their scope. As a consequence, the funcion searchs outside the function's environment (or local environment) and goes one level up and catch *z* in the global environment. In case *z* is defined it clearly show the output *(1 ,2 ,15)*, otherwise it prints an error. For example (uncomment to see the mistake):

```{r, echo=TRUE, eval=TRUE}
# z <- 15
# f <- function() {
#  x <- 1
#  y <- 2
#  c(x, y, z, d)
# }
# f()
```

See that variable **d** is not defined in the local or global environment. For this reason, it shows an error as follow: *Error in f() : object 'd' not found*.

Okay, now I open a new question: What happen wheter a variable is defined inside and outside the function scope?. Let see

```{r, echo=TRUE, eval=TRUE}
z <- 15
producto <- function(z) {
 z <- 1
 y <- 2
 z*y
}
producto()
```
The answer is simple, the function prioritize the local variable over the global variable.




# Module 2:  Functional Programming (FP)

Several times we work with repeated procedures that need to be automatized. In order to do this, we requiere the use **for** loops. The **For** loop is a finite repeated procedure to couldn't help write the same pice of code for different values. An example as following:

- Calculation of the media **col_media** 

```{r, echo=TRUE, eval=TRUE}
#The st4gi library
library(st4gi)
#By now, we are going to extract the first two dull columns.
df <- pjpz09[,-c(1:2)]

#The pre-allocation of 'med' variable that storage the medians for each column or trait
col_media <- vector("numeric", length = ncol(df))
#A second option is using: 

#seq_along is the same as write  'i in 1:length(col_media)' 
for(i in seq_along(col_media) ){
  col_media[i] <- mean(df[[i]])
  #col_media[i] <- mean(df[,i])
}
col_media
```


- Calculation of the median **col_mediana** 

```{r, echo=TRUE, eval=TRUE}
#The st4gi library
library(st4gi)
df <- pjpz09[,-c(1:2)]
col_mediana <- vector("numeric", length = ncol(df))
for(i in seq_along(col_media) ){
  col_mediana[i] <- median(df[[i]])
}
col_media
```

- Finally the standard deviation **col_devs**

```{r, echo=TRUE, eval=TRUE}
#The st4gi library
library(st4gi)
df <- pjpz09[,-c(1:2)]
col_devs <- vector("numeric", length = ncol(df))
for(i in seq_along(col_devs) ){
  col_devs[i] <- sd(df[[i]])
}
col_devs
```

As we can see, the for loop help us to calculate iteratively the mean, median and s.deviation for different *i* columns. However a new question arises based on this examples: *How about iterate functions and not only values?*.  Following the first question derives the second one, *Functions could be arguments of other functions?*

To elucidate these doubts, we are going to pass a function as argument of other function called **col_summary**. **col_summary** has two arguments **df** and **fun**, where *df* is the data and *fun* is a function argument.

```{r, echo=TRUE, eval=TRUE}
library(st4gi)
df <- pjpz09[,-c(1:2)]

col_summary <- function(df, fun) {
    #Second way to pre-allocate, using numeric(length(vector or dataframe))  
    output <- numeric(length(df))
    for (i in seq_along(df)) {
        output[i] <- fun(df[[i]])
    }
    output
}
#Let make some calculations over the columns using the mean, median and standard desviation
col_summary(df, fun = mean)
col_summary(df, fun = median)
col_summary(df, fun = sd)
```

It is plain to see that a function can be an argument of other function. Note that the *fun* argument is used without double quotes. 

The R-base program provides function which use the functional programming approach. The set of function of this family are: apply, sapply, rapply, etc. Lets review an example:

```{r, echo=TRUE, eval=TRUE}
### Using sapply function ###
library(st4gi)
df <- pjpz09[,-c(1:2)]
sapply(df, mean)
````

```{r, echo=TRUE, eval=TRUE}
### Using our implentation
col_summary(df, fun = mean)
````


Along the year, many r-packages has rised and one of them tackle the functional programming approach, this is the **purrr** package from Hadley Wickham. **purr** package (written with three *r*s similarly to cat's meowth) provides a robust implementation of the FP paradigm in R. Besides, it wrapps *C++* code to make the computing much more faster when you are dealing with big datasets or the famous *Big Data*.


## The purr package

To begin with *purrr* package, we will use the set of *map* functions. For example, lets calculate the mean over all traits using *purr*

```{r, echo=TRUE, eval=TRUE}
library(purrr)
map_dbl(.x = df, .f= mean)
````

What *purrr* basically does is:

1. Loop over a vector *.x*. More generally, *.x* could be a data structue as vectors, data frames, lists, etc.
2. Do something to each element *.f*. For instance, it pass the mean function over all the data structure.
3. Return the results.


The map function varies in their return type. Lets see:

- *map()* returns a list
- *map_dbl()* returns a double vector
- *map_lgl()* returns a logical vector
- *map_int()* returns a integer vector
- *map_chr()* returns a character vector
